---
title: "Final Project Code"
author: "Grace"
date: "12/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load Libraries required for data analysis
library(tidyverse)
library(here)
```

# Load and Tidy Data

```{r message=FALSE}
# Load region CO2 flux dataframe
data <- read_csv(here("data/regional_survey.csv"))

# The columns of interest are AirTemp_C, and Fmmol_mmol_m2_day
# Carbon flux units are CO2 flux in mmol per m2 per day

# Tidy region CO2 flux dataframe
data_tidy <- data %>% mutate(Habitat = tolower(Habitat)) %>% # Change all Habitat names to lowercase
  rename(c_flux = Fmmol_mmol_m2_day, # Rename key variables 
         air_temp = AirTemp_C) %>%
  filter(air_temp >-100) %>% filter(c_flux >-100) %>% # filter out -9999 missing values for air temp and C flux
  filter(Habitat != "tk-pond") %>% # remove tk-pond observations due to few (7) observations 
  filter(Habitat != "ocean") #remove ocean due to all equivalent air temperature values and time
```

## Diagnostic Plots

```{r}
# Frequency plot of C Flux
ggplot(data = data_tidy, aes(x=c_flux))+
  geom_freqpoly() + theme_bw()+
  labs(x = expression(Carbon ~ Flux ~ (mmol ~ '/' ~ m^{2} ~ '/' ~ day)))
```

# All-Habitat Correlation

This plot, correlation test and linear model is in order to determine the relationship between air temperature and carbon flux across all aquatic habitat types observed within the dataset. 

```{r}
# Quantify the overall Carbon flux average (across all habitat types)
mean(data_tidy$c_flux)
sd(data_tidy$c_flux)

# Plot overall C Flux ~ Air Temperature
ggplot(data = data_tidy, mapping = aes(x= air_temp, y=(c_flux))) +
  geom_point() + geom_smooth(method = lm) + theme_bw()+
  labs(x="Air Temperature (°C)", y = expression(Carbon ~ Flux ~ (mmol ~ '/' ~ m^{2} ~ '/' ~ day)))

# Pearson's Correlation Test
cor.test(data_tidy$c_flux, data_tidy$air_temp)

# Linear model (y~x) 
lm_overall <- lm(c_flux ~ air_temp, data = data_tidy)
summary(lm_overall)
``` 

## Individual Habitat Carbon flux

This first boxplot shows the average and spread of carbon flux data separated by habitat. I then run an ANOVA to test for significant differences between carbon flux at different habitat types.
I then run a Tukey HSD test test on the estimated ANOVA model to identify groups of habitats that are significantly different from each other with regard to carbon flux. 

```{r}
# C flux Boxplot by Habitat
ggplot()+
  geom_boxplot(data = data_tidy, mapping = aes(x= air_temp, y=(c_flux), color = Habitat)) +
  theme_bw() +
  labs(x="Habitat", y = expression(Carbon ~ Flux ~ (mmol ~ '/' ~ m^{2} ~ '/' ~ day))) 

# ANOVA, test for statistically significant difference between carbon flux by habitat
aov_cflux_hab <- aov(c_flux ~ Habitat, data = data_tidy)
summary(aov_cflux_hab) 

# Run Tukey HSD test on estimated ANOVA model
TukeyHSD(aov_cflux_hab)

# Show Tukey HSD test output
library(agricolae) # load package that helps to view output
aov_cflux_hab <- HSD.test(aov_cflux_hab, trt = "Habitat", alpha = 0.05) # set treatment variable to Habitat and p-value cutoff to 0.05
aov_cflux_hab # view output

```

## Analyze the relationship between air temperature and carbon flux by habitat

```{r}
# C Flux ~ Air temp by Habitat
ggplot() +
    geom_point(data = data_tidy, mapping = aes(x= air_temp, y=(c_flux), color = Habitat)) + 
    geom_smooth(data = data_tidy, mapping = aes(x= air_temp, y=(c_flux), color = Habitat), 
    method = lm, se =F) + theme_bw() +
  labs(x="Air Temperature (°C)", y = expression(Carbon ~ Flux ~ (mmol ~ '/' ~ m^{2} ~ '/' ~ day)))

```

```{r}
unique(data_tidy$Habitat) # list all unique habitat types
 
## Create discrete data frames for each habitat type
pond <- data_tidy %>% filter(Habitat == "pond")
lake <- data_tidy %>% filter(Habitat == "lake")
lagoon <- data_tidy %>% filter(Habitat == "lagoon")
river <- data_tidy %>% filter(Habitat == "river")
salt_river <- data_tidy %>% filter(Habitat == "salt-river")

# Correlation test for each habitat type (* denotes significant p value for correlation)
cor.test(pond$c_flux, pond$air_temp) #p-value = 0.01328 *
cor.test(lake$c_flux, lake$air_temp) #p-value = 0.5811
cor.test(lagoon$c_flux, lagoon$air_temp) #p-value = 1.613e-13 *
cor.test(river$c_flux, river$air_temp) #p-value = 0.8745
cor.test(salt_river$c_flux, salt_river$air_temp) #p-value = 0.05199 *?

# Linear Regression for significantly correlated habitat data (pond, lagoon and salt river)
pond_lm <- lm(c_flux ~ air_temp, data = pond)
lagoon_lm <- lm(c_flux ~ air_temp, data = lagoon)
saltriver_lm <- lm(c_flux ~ air_temp, data = salt_river)

## View summary data from linear models
summary(pond_lm)
summary(lagoon_lm)
summary(saltriver_lm)
```

